Sequential rank agreement methods for comparison of ranked lists

### 2016-07-19 latex/superranker-manuscript.tex
small fixes

\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\Cov}{Cov}
\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}

@@

\begin{theorem}
  Assume that $\{R_l(X)\}_{l=1}^L$ are independent draws from a
  probability distribution $Q$ on the set of lists $\Pi$. Then,
  $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty =
  o_P(1)$.
\label{thm:consistency}
\end{theorem}
\proof See Appendix \ref{sec:appA}.

@@

The consistency of the estimated change-point, $\widehat{d^\ast_L}(\widehat{q}_L)$, follows from Theorem \ref{thm:consistency} by the following corollary. 
\begin{corollary}\label{thecor}
Let $\widehat{q}_L$ be a threshold function such that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ for some limiting
function $q$.  Then, $\widehat{d^\ast_L}(\widehat{q}_L) \overset{P}{\longrightarrow} d^\ast(q)$ for $L \rightarrow \infty$.
\label{cor:consistency}
\end{corollary}
\proof See Appendix \ref{sec:appB}.

@@

We start by defining the following cadlag function
\begin{align}
  N(p; d) = \sum_{s=1}^p 1\left(Q(R(X_s) \leq d) > 0\right), \quad p = 1,\ldots, P
\end{align}
which runs through the list elements in an arbitrary order and counts
how many list elements have a strictly positive probability to have
rank less than or equal to $d$ under the probability measure $Q$.
Note that the cardinality of the set $S(d)$ in equation
(\ref{eq:sumset}) is equal to the counting process evaluated at the last element: $|S(d)| = N(P;d)$.
The empirical counterpart of the counting process is given by
\begin{align}
  \widehat{N}_L(p; d) &= \sum_{s=1}^p 1\left( \frac{1}{L} \sum_{l=1}^L 1(R_l(X_s) \leq d) > 0\right)\\
        &= \sum_{s=1}^p 1\left(\widehat{Q}_L(R(X_s) \leq d) > 0\right)\nn.
\end{align}
The joint law of the sequence
$\{\widehat{N}_L(p;d)\}_{1\leq p \leq P}$ is completely determined by
the finite set of the jump times of $N(p;d)$. Since
$\{1(R_l(X_s) \leq d); l=1,\dots, L\}$ consists of independent and
identically Bernoulli distributed variables with expectation
$Q(R(X_s) \leq d)$ it follows from the law of large numbers that
$\widehat{Q}_L(R(X_p) \leq d) \overset{P}{\longrightarrow} Q(R(X_p)
\leq d)$ for every $p$ and $d$ as $L \rightarrow \infty$. 
Therefore, we have
\begin{align}
  \sup_{d \in 1,\ldots,P}\sup_{p \in 1,\ldots, P} \left|\widehat{N}_L(p; d) - N(p; d)\right| = o_P(1).\label{eq:countingConv}
\end{align}

@@

\section{Proof of Corollary \ref{cor:consistency}}
\label{sec:appB}

Recall that $\mathcal{L}$ is the superlevel set of list positions
where the sequential rank agreement exceeds the threshold function
$q$. Let $A \triangle B = (A \setminus B) \cup (B \setminus A)$ is the
symmetric difference between sets $A$ and $B$. It is then sufficient
to show that
$|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q)|$
converges to zero in probability as $L \rightarrow \infty$. By the
triangle inequality we have that
\begin{align}
 \left|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q)\right| &= \left|\left(\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \widehat{\mathcal{L}}_L(q)\right) \cup \left(\widehat{\mathcal{L}}_L(q) \triangle \mathcal{L}(q)\right)\right|\label{eq:corSetDiff}\\
   &\leq \left|\widehat{\mathcal{L}}_L(q) \triangle \mathcal{L}(q)\right| + \left|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \widehat{\mathcal{L}}_L(q)\right|\nn
\end{align}
Looking at the first term we have
\begin{align}
  \left|\widehat{\mathcal{L}}_L(q) \triangle \mathcal{L}(q)\right| &= \sum_{d=1}^P \left|1\left(\widehat{\textrm{sra}}_L(d) \geq q(d)\right)  - 1\left(\textrm{sra}(d)  \geq q(d)\right)\right|\\
   &=
\end{align}
Similarly we have
\begin{align}
\left|\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \widehat{\mathcal{L}}_L(q)\right| &= \sum_{d=1}^P \left|1\left(\widehat{\textrm{sra}}_L(d) \geq \widehat{q}_L(d)\right) - 1\left(\widehat{\textrm{sra}}_L(d) \geq q(d)\right)\right|\\
   &=
\end{align}

%We have that 	
%\begin{align}
%  |\widehat{d^\ast_L}(\widehat{q}_L) - d^\ast(q)| &\leq  |\widehat{d^\ast_L}(q)  - d^\ast(q)| + |\widehat{d^\ast_L}(\widehat{q}_L)  - \widehat{d^\ast_L}(q_L)|
%\end{align}
%The first term is of order $o_P(1)$ by the continuous mapping theorem if a) $|\inf \widehat{\mathcal{L}}_L(q) - \inf  \mathcal{L}(q)| = o_P(1)$ and b) $|1(|\widehat{\mathcal{L}}_L(q)| = 0) - 1(|\mathcal{L}(q)| = 0)| = o_P(1)$ since $1(|\mathcal{L}(q)| > 0) = 1 - 1(|\mathcal{L}(q)| = 0)$. Similarly, the second term is of order $o_P(1)$ by the same reasoning if c) $|\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) - \inf \widehat{\mathcal{L}}_L(q)| = o_P(1)$ and d) $|1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0) - 1(|\widehat{\mathcal{L}}_L(q)| = 0)| = o_P(1)$.

This completes the proof.

>>>>

\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\Cov}{Cov}
\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
\newcommand{\plim}{\operatornamewithlimits{plim}}

@@

\begin{theorem}
  Assume that $\{R_l(X)\}_{l=1}^L$ are independent draws from a
  probability distribution $Q$ on the set of lists $\Pi$. Then,
  $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty =
  o_P(1)$.
\label{thm:consistency}
\begin{proof}
See Appendix \ref{sec:appA}.
\end{proof}
\end{theorem}

@@

The consistency of the estimated change-point, $\widehat{d^\ast_L}(\widehat{q}_L)$, follows from Theorem \ref{thm:consistency} by the following corollary. 
\begin{corollary}\label{thecor}
Let $\widehat{q}_L$ be a positive threshold function such that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ for some limiting
function $q$.  Then, $\widehat{d^\ast_L}(\widehat{q}_L) \overset{P}{\longrightarrow} d^\ast(q)$ for $L \rightarrow \infty$.
\label{cor:consistency}
\begin{proof}
See Appendix \ref{sec:appB}.
\end{proof}
\end{corollary}

@@

We start by defining the following cadlag function
\begin{align}
  N(p; d) = \sum_{s=1}^p 1\left(Q(R(X_s) \leq d) > 0\right), \quad p = 1,\ldots, P
\end{align}
which runs through the list elements in an arbitrary order and counts
how many list elements have a strictly positive probability to have
rank less than or equal to $d$ under the probability measure $Q$.
Note that the cardinality of the set $S(d)$ in equation
(\ref{eq:sumset}) is equal to the counting process evaluated at the last element: $|S(d)| = N(P;d)$.
The empirical counterpart of the counting process is given by
\begin{align}
  \widehat{N}_L(p; d) &= \sum_{s=1}^p 1\left( \frac{1}{L} \sum_{l=1}^L 1(R_l(X_s) \leq d) > 0\right)\\
        &= \sum_{s=1}^p 1\left(\widehat{Q}_L(R(X_s) \leq d) > 0\right)\nn.
\end{align}
The joint law of the sequence
$\{\widehat{N}_L(p;d)\}_{1\leq p \leq P}$ is completely determined by
the finite set of the jump times of $N(p;d)$. Since
$\{1(R_l(X_s) \leq d); l=1,\dots, L\}$ consists of independent and
identically Bernoulli distributed variables with expectation
$Q(R(X_s) \leq d)$ and finite variance it follows from the law of large numbers that
$\widehat{Q}_L(R(X_p) \leq d) \overset{P}{\longrightarrow} Q(R(X_p)
\leq d)$ for every $p$ and $d$ as $L \rightarrow \infty$. 
Therefore, we have
\begin{align}
  \sup_{d \in 1,\ldots,P}\sup_{p \in 1,\ldots, P} \left|\widehat{N}_L(p; d) - N(p; d)\right| = o_P(1).\label{eq:countingConv}
\end{align}

@@

\section{Proof of Corollary \ref{cor:consistency}}
\label{sec:appB}
Recall that $\mathcal{L}$ is the superlevel set of list positions
where the sequential rank agreement exceeds the threshold function
$q$. Let $A \triangle B = (A \setminus B) \cup (B \setminus A)$ be the
symmetric difference between sets $A$ and $B$. It is then sufficient
to show that $\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q) \overset{P}{\longrightarrow} \emptyset$
for $L \rightarrow \infty$ where $\emptyset$ denotes the empty set.
We have that
\begin{align}
  \plim_{L \rightarrow \infty}\left(\widehat{\mathcal{L}}_L(\widehat{q}_L) \triangle \mathcal{L}(q)\right) &=   \left(\plim_{L \rightarrow \infty}\widehat{\mathcal{L}}_L(\widehat{q}_L)\right) \triangle \mathcal{L}(q)\\
  &= \left\{d : \plim_{L\rightarrow \infty} \left(\widehat{\textrm{sra}}_L(d) - \widehat{q}_L(d)\right)\geq 0\right\} \triangle \mathcal{L}(q)\nn\\
  &= \left\{d : \textrm{sra}(d) - q(d) \geq 0\right\}\triangle \mathcal{L}(q)\nn\\
  &= \mathcal{L}(q)\triangle \mathcal{L}(q)\nn\\
  &= \emptyset\nn
\end{align}
as a consequence of Theorem \ref{thm:consistency}, the assumption that $\left\|\widehat{q}_L - q\right\|_\infty = o_P(1)$ and the continuous mapping theorem. This completes the proof.
